Exercise 3: Software Architecture in Practice Case: Beer Web Store

### 2008-03-17 trunk/SAiP/Module1/H3/doc/softwarearchitecture.tex
Fixing spelling error and Figure references

A new Beer Web Store for specialty beers has the following main use
case:
\begin{quote}
  The customer browses the beer catalogue for descriptions and
  reviews. The customer selects appropriate beers for purchase and
  these are added to his shopping basket. When the customer checks
  out, payment is validated and he is presented with a receipt
\end{quote}
As a first step, the software architect of this systems has decided to
use a layered architectural style (with clients, server, and a
database in separate layers). He has decided that the communication to
the database will be done via a Fa\c cade pattern. The system will be
built in Java.

@@

\subsubsection{Choose the module to decompose}
As this is the beginning of the architectual design then the module is the entire system. In later iterations the sub-components (the layers, the components in the layers, etc.) will be decomposed.

@@

\begin{description}
    \item[a. Choose the architectual drivers]
  Here the quality attributes of the system would be defined by the stakeholders of the system and then a vote would be used to determine which of the quality attributes are architectual drivers. The quality attributes would be expressed as quality attribute scenarios.
    \item[b. Choose an architectual pattern]
  Based on the architectual drivers selected in 1, use the mapping in Bass et al. as well as other literature and experience to determine some tactics for achieving the architectual drivers for the system. Then select some architectual patterns to realize the tactics.
    \item[c. Instantiate modules using multiple views]
  Based on the tactics and patterns found above the architecture for the system may be expressed using UML. It is \emph{always} required to use multiple views in order to properly describe the architecture. A good approach is the "3 + 1" structure, which uses three viewpoints; Module viewpoint, Component \& Connector viewpoint and Allocation viewpoint. These three viewpoint is expressed in UML using Package diagrams, Object and sequence diagrams, and deployment diagrams respectively. These diagrams should in the first iteration only look at relations and interactions between the main components (the layers).
    \item[d. Define interfaces]
  Define the interfaces between the main components (the layers).
    \item[e. Verify scenarios and use cases]
  Make sure that the steps a - d covered all the parts of the scenarios. If not include the missing parts and updates the scenarios/architecture is inconsistencies are found.
\end{description}
\subsubsection{Repeat the steps above for every module that needs further decomposition}
The first iteration only look at the overall architecture (the leation and communication between layers). The following iterations will decompose the individual layers and as needed the individual components in the layers.

@@

\subsubsection{Relations}
The figure xx shows some of the relations, namely the statical, yet it is also important to know the dynamical behaviour. These may be expressed by e.g. a sequence diagram, as shown in Figure yy. In these it may be seen that the Client and Server has a week semantic relation through the http protocol, and the Server and Database has a strong syntactic relation from a compiled compatibility.

@@

Perry and Wolf suggest seperating the model of the architecture into three components, as mentioned in the question. The components relate in the sense that the \emph{elements} has a particular \emph{form} and them and their relations are chosen based on a \emph{rational}.

There are three class of architectual elements defined.
\begin{itemize}
    \item[Processing elements] Transforms data elements.
    \item[Data elements] Information to be transformed or used.
    \item[Connecting elements] The \emph{glue} that hold the architecture together. May at times be \emph{Data elements}, \emph{Connecting elements} or both. The connecting elements are the relations between the different elements, e.g. procedure calls or message parsing.
\end{itemize}

The architectual form describe the architectual properties and the relationships and their relative importance. This is important as some architectual properties
may be vital to the design and others may be gold-plating. These properties are used to put constraints on the elements and also define the relative importanse of these constraints. The relationships are similar to the properties except they are used to put constraints and relative priority on the relations between elements.

A software architecture is a selection and description of architectual choices. These choices are, however, difficult to convay without including the reason for the choices made, for the elemenents and form not chosen and for choices not made due to lack of information or because they were deemed irrelevant.

@@

\begin{itemize}
    \item Processing elements.
	 \begin{itemize}
	    \item[Client] Receive user entries and process them into http requests
	    \item[Server] Receive http requests and transform them into high level database lookup
	    \item[DatabaseFacade] Transform high level database lookup into proprietary database lookup
	 \end{itemize}
    \item Data elements.
	 \begin{itemize}
	    \item[Request data] The packets of data flowing from the client to the server and form the server to the database, e.g. Beer type, Credit card information, search refinement parameter, etc.
	 \end{itemize}
    \item Connecting elements.
	 \begin{itemize}
	    \item[HTTP] Protocol for connecting the server and client (loose coupling).
	    \item[Database facade methods] Protocol for connecting the server to the database (medium-loose coupling).
	 \end{itemize}
\end{itemize}

\subsubsection{Form in Beer Web Store}
\begin{itemize}
    \item[Properties] The Server and the database, hidden behind the Facade, is the backbone of the Beer Web Store. If either is unavailable then the other is irrelevant. Multiple instances of both Server and Database may be running simultaniously in order to supply hot-swapping. The importance of up-time (availability), security (payment), usability (transmission abilities), etc. must be defined here.
    \item[Relationship] The Client may only see the server through its HTTP interface, and the Server may only access the database through the Database Facade. Other demands to the relationship (certain commands to the database may only be accessed when using https).
\end{itemize}

@@

If the references to software and computing is removed the definition talks about elements, structures and properties, and their relations. This definition may be used about anything which may be broken down into generalised components. As an example could be used the construction of a rifle. Describing the elements that goes into making a rifle (hammer, barrol, ...), the structures that comes from putting them together, and the individual relationships that must be met (form fit, etc.). The elements may fit together in different ways and multiple structures may be required to complete the description. (the hammer up, the hammer down, etc.). The properties are in fact only the externally visible to the given element, as those are the only properties that are relevant to the architecture. These details include things like the strength of the metal, which may be different for the hammer and the pipe, the price range, produktion time, etc. while things like the cheak-pad may be optional.

The description of an architecture, as defined by Bass et. al. is therefore very general and apart from the use of words like "software" and "computer", may apply to any architecture, be it a building, a produktion line for a rifle, or a embedded software system.

@@

There are also two reasons for creating an architecture; in order to create a product that exhibits some of the quality attributes defined in the architecture (modifiability, usability, ...) and in order to reuse the architecture (or part of the architecture) as a product line for creating other products. This latter reason is described in section yy.

Creating an architecture up front allows for the definition of all properties of the architecture; the quality attributes and their individual priorities, the individual elements/components and how they interconnect both at runtime and statically, etc. This has the benifit that when the design and implementation phase is begun it may easily be broken down into parallel activities and integrated later, as everyone has a clear picture of what is important and what the interfaces are.

Unfortunately defining the architecture fully up front only makes sense if there is enough information in advance to create the architecture form. Very often there are "dark horses", things that do not become apparent until the design and implementation phase. If it is not believed that sufficient information exist to fully define the architecture, then doing so may not only be a waste, it may lead to the very dangerous loose-loose principle, where e.g. an overhead is introduced to service modifiability, only to be hacked in order to service a performance issue that was not apparent at the architectual design pahse. This means that part of the overhead is preserved (in memory-footprint if nowhere else), without preserving any of the modifiability advantages.

Another requirement to defining the architecture fully up front has to do with the organization. It may be good and dandy to create a fully defined architecture with focus on parallel implementation, if there is only one in-house developer to implement the product.
Based on this the Beer Web Store architect must have decided that there are sufficient information to fully define the architecture and also that there is a reason (parallel implementation, out-sourceing of the design/implementation, etc.).

Asuming that the architect in question has followed the course Software Architecture in Practice, he has obviously looked at the architectual properties, the stake holders, the organization and the requirement and deemed that there is sufficient grounds to fully specify the architecture in advance, with the risks and possible gains that it has.

>>>>

A new Beer Web Store for speciality beers has the following main use
case:
\begin{quote}
  The customer browses the beer catalogue for descriptions and
  reviews. The customer selects appropriate beers for purchase and
  these are added to his shopping basket. When the customer checks
  out, payment is validated and he is presented with a receipt
\end{quote}
As a first step, the software architect of this systems has decided to
use a layered architectural style (with clients, server, and a
database in separate layers). He has decided that the communication to
the database will be done via a Fa\c cade pattern. The system will be
built in Java.

@@

\subsubsection{Choose the module to decompose}
As this is the beginning of the architectural design then the module is the entire system. In later iterations the sub-components (the layers, the components in the layers, etc.) will be decomposed.

@@

\begin{description}
    \item[a. Choose the architectural drivers]
  Here the quality attributes of the system would be defined by the stakeholders of the system and then a vote would be used to determine which of the quality attributes are architectural drivers. The quality attributes would be expressed as quality attribute scenarios.
    \item[b. Choose an architectural pattern]
  Based on the architectural drivers selected in 1, use the mapping in Bass et al. as well as other literature and experience to determine some tactics for achieving the architectural drivers for the system. Then select some architectural patterns to realize the tactics.
    \item[c. Instantiate modules using multiple views]
  Based on the tactics and patterns found above the architecture for the system may be expressed using UML. It is \emph{always} required to use multiple views in order to properly describe the architecture. A good approach is the "3 + 1" structure, which uses three viewpoints; Module viewpoint, Component \& Connector viewpoint and Allocation viewpoint. These three viewpoints are expressed in UML using Package diagrams, Object and sequence diagrams, and deployment diagrams respectively. These diagrams should in the first iteration only look at relations and interactions between the main components (the layers).
    \item[d. Define interfaces]
  Define the interfaces between the main components (the layers).
    \item[e. Verify scenarios and use cases]
  Make sure that the steps a - d covered all the parts of the scenarios. If not include the missing parts and updates the scenarios/architecture is inconsistencies are found.
\end{description}
\subsubsection{Repeat the steps above for every module that needs further decomposition}
The first iteration only look at the overall architecture (the location and communication between layers). The following iterations will decompose the individual layers and as needed the individual components in the layers.

@@

\subsubsection{Relations}
The figure \ref{fig:module_view} shows some of the relations, namely the statical, yet it is also important to know the dynamical behaviour. These may be expressed by e.g. a sequence diagram, as shown in Figure \ref{fig:sequence}. In these it may be seen that the Client and Server has a week semantic relation through the http protocol, and the Server and Database has a strong syntactic relation from a compiled compatibility.

@@

Perry and Wolf suggest separating the model of the architecture into three components, as mentioned in the question. The components relate in the sense that the \emph{elements} has a particular \emph{form} and them and their relations are chosen based on a \emph{rational}.

There are three class of architectural elements defined.
\begin{itemize}
    \item[Processing elements] Transforms data elements.
    \item[Data elements] Information to be transformed or used.
    \item[Connecting elements] The \emph{glue} that hold the architecture together. May at times be \emph{Data elements}, \emph{Connecting elements} or both. The connecting elements are the relations between the different elements, e.g. procedure calls or message parsing.
\end{itemize}

The architectural form describe the architectural properties and the relationships and their relative importance. This is important as some architectural properties
may be vital to the design and others may be gold-plating. These properties are used to put constraints on the elements and also define the relative importance of these constraints. The relationships are similar to the properties except they are used to put constraints and relative priority on the relations between elements.

A software architecture is a selection and description of architectural choices. These choices are, however, difficult to convey without including the reason for the choices made, for the elements and form not chosen and for choices not made due to lack of information or because they were deemed irrelevant.

@@

\begin{itemize}
    \item Processing elements.
	 \begin{itemize}
	    \item[Client] Receive user entries and process them into http requests
	    \item[Server] Receive http requests and transform them into high level database lookup
	    \item[DatabaseFacade] Transform high level database lookup into proprietary database lookup
	 \end{itemize}
    \item Data elements.
	 \begin{itemize}
	    \item[Request data] The packets of data flowing from the client to the server and form the server to the database, e.g. Beer type, Credit card information, search refinement parameter, etc.
	 \end{itemize}
    \item Connecting elements.
	 \begin{itemize}
	    \item[HTTP] Protocol for connecting the server and client (loose coupling).
	    \item[Database fa\c ade methods] Protocol for connecting the server to the database (medium-loose coupling).
	 \end{itemize}
\end{itemize}

\subsubsection{Form in Beer Web Store}
\begin{itemize}
    \item[Properties] The Server and the database, hidden behind the Fa\c ade, is the backbone of the Beer Web Store. If either is unavailable then the other is irrelevant. Multiple instances of both Server and Database may be running simultaneously in order to supply hot-swapping. The importance of up-time (availability), security (payment), usability (transmission abilities), etc. must be defined here.
    \item[Relationship] The Client may only see the server through its
 HTTP interface, and the Server may only access the database through the Database Fa\c ade. Other demands to the relationship (certain commands to the database may only be accessed when using https).
\end{itemize}

@@

If the references to software and computing is removed the definition talks about elements, structures and properties, and their relations. This definition may be used about anything which may be broken down into generalised components. As an example could be used the construction of a rifle. Describing the elements that goes into making a rifle (hammer, barrol, ...), the structures that comes from putting them together, and the individual relationships that must be met (form fit, etc.). The elements may fit together in different ways and multiple structures may be required to complete the description. (the hammer up, the hammer down, etc.). The properties are in fact only the externally visible to the given element, as those are the only properties that are relevant to the architecture. These details include things like the strength of the metal, which may be different for the hammer and the pipe, the price range, production time, etc. while things like the cheek-pad may be optional.

The description of an architecture, as defined by Bass et. al. is therefore very general and apart from the use of words like "software" and "computer", may apply to any architecture, be it a building, a production line for a rifle, or a embedded software system.

@@

There are also two reasons for creating an architecture; in order to create a product that exhibits some of the quality attributes defined in the architecture (modifiability, usability, ...) and in order to reuse the architecture (or part of the architecture) as a product line for creating other products. This latter reason is described in section \ref{sec:pl_q2}.

Creating an architecture up front allows for the definition of all properties of the architecture; the quality attributes and their individual priorities, the individual elements/components and how they interconnect both at runtime and statically, etc. This has the benefit that when the design and implementation phase is begun it may easily be broken down into parallel activities and integrated later, as everyone has a clear picture of what is important and what the interfaces are.

Unfortunately defining the architecture fully up front only makes sense if there is enough information in advance to create the architecture form. Very often there are "dark horses", things that do not become apparent until the design and implementation phase. If it is not believed that sufficient information exist to fully define the architecture, then doing so may not only be a waste, it may lead to the very dangerous loose-loose principle, where e.g. an overhead is introduced to service modifiability, only to be hacked in order to service a performance issue that was not apparent at the architectural design phase. This means that part of the overhead is preserved (in memory-footprint if nowhere else), without preserving any of the modifiability advantages.

Another requirement to defining the architecture fully up front has to do with the organization. It may be good and dandy to create a fully defined architecture with focus on parallel implementation, if there is only one in-house developer to implement the product.
Based on this the Beer Web Store architect must have decided that there are sufficient information to fully define the architecture and also that there is a reason (parallel implementation, out-sourcing of the design/implementation, etc.).

Assuming that the architect in question has followed the course Software Architecture in Practice, he has obviously looked at the architectural properties, the stake holders, the organization and the requirement and deemed that there is sufficient grounds to fully specify the architecture in advance, with the risks and possible gains that it has.
